---
title: "Racism Within the United States Prison System"
author: "Cole Meier"
date: '2022-11-01'
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r echo=FALSE, eval=TRUE, message=FALSE}
source("~/Documents/Info_201/Assignments/a4-coletmeier2/source/a4-helpers.R")
source("~/Documents/Info_201/Assignments/a4-coletmeier2/source/analysis.R")
```

## Abstract

This analysis is examining the United States prison system, and the
potential injustices within it. The data showed that they did not keep
record of race in the prison system until the 80's. Once they began
keeping track of race, a definite racial bias appeared in the data. The
proportion of incarcerated black people far out weighed their total
general population proportion.

## Introduction

The United States has a history of repeated and targeted oppression
towards black people in America. This report will specifically dissect
the in proportionate incarceration of different races in America. This
will be viewing whether there is a direct correlation between race and
the likelihood of incarceration in this country. The direct stakeholders
for this cause are the millions of minorities in this country that are
directly affected by the findings of this report. This has many outside
implications, including the amount of single mother homes, generational
wealth, and biases that effect the well being of individuals. The goal
of this report is to educate the masses on the racism that is ingrained
in our society for an area that is often overlooked. This will
optimistically aid in chipping away at the matrix of domination in this
country. This goal will be achieved by answering the following:

1.  What percentage of the prison population is a minority?
2.  Which races has the highest proportion of incarceration per capita?
3.  Which race has the lowest proportion of incarceration per captia?

Answering thesequestions will enable us to make a complete evaluation of
the biases within the prison system in America.

## Data Summary

The proportion of the total number of black people incarcerated is much
higher compared to how much of the population they make up. Black people
make up around 13% of the population, but they make up on average 27% of
the total prison population. Compared to white population which makes up
around 57% of the total population, but only make up on average 45
percent of the total prison population. This points to possible racial
bias in the American prison system. Another important statistic is the
top-5 states with the highest incarceration. Most of the states on the
list are more expected due to their large populations, however, Georgia
is outside the top 5 highest total population, but is number 4 in total
incarcerated population. This shows the possibility of an unfair
judicial system and over policing in Georgia.

## Growth of the U.S. Prison Population

This bar graph is a representation of the total jail populations from
1970 to 2018. This shows that the number of people in jail was steady
through the seventies. Then is started rapidly raising until the peak in
2008. Since 2008, the total jail population has remained relatively
consistent.

```{r echo=FALSE}
plot_jail_pop_for_us()
```

The total prison population by year

## Growth of Prison Population by State

This bar graph is a representation of the total jail population of
Washington, California, and Oregon, from 1970 to 2018. This has a very
similar trend to the total jail population of the whole country,
however, the upwards trend is a lot more steep. This upwards trend in
total jail population happens between 1980 and the early 2000's.

```{r echo=FALSE}
plot_jail_pop_by_states(c("WA", "CA", "OR"))
```

Total jail population of Washington, California, and Oregon, from 1970
to 2018.

## Growth of Black Prison Population

This bar graph is a representation of the growth of the total black jail
population from 1970 to 2018. This graph shows that they did not keep
track of the quantity of black people in prisons until 1985, causing the
large spike. After they began to get recorded, the amount of black
people incarcerated skyrocketed. The number began to decrease in the mid
2000's, and began to level out after around 2013.

```{r echo=FALSE}
plot_black_jail_pop()
```

Total jail population of black people in America

## Area Map of Percentage of Black Prison Population

This is the representation of the two cities with the highest amount of
black people incarcerated in 2018. These locations can signify how the
injustice of black people in prison can become normalized within the
culture. These areas most likely also have higher numbers of black
families without parental figures in the home because of this injustice.

```{r echo=FALSE}
plotting_cities()
```

The cities with the highest number of incarcerated black people in 2018
